---
title: "2025 MLB Postseason Statcast Analysis"
author: "Tung-Tsan WU"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: flatly
---

```{r setup, include=FALSE}
# Basic setup: hide warning and message outputs in the final HTML report
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 1. Data Import & Cleaning

First, we load the necessary packages and the MLB Statcast dataset, followed by data formatting and cleaning to prepare for analysis.

```{r data-prep}
library(tidyverse)
library(lubridate)
library(psych)
library(knitr)

# Load data
df_mlb <- read_csv("Data_MLB_2025_StatcastPostseason_PitchByPitch_20251102a.csv",
                   show_col_types = FALSE)

# Data cleaning
clean_df <- df_mlb %>%
  select(
    -spin_dir, -spin_rate_deprecated, -break_angle_deprecated,
    -break_length_deprecated, -tfs_deprecated, -tfs_zulu_deprecated,
    -umpire, -sv_id
  ) %>%
  select(-matches("fielder_2\\.\\d+")) %>%
  mutate(
    game_date = ymd(game_date),
    across(c(pitch_type, pitch_name, description, events,
             if_fielding_alignment, of_fielding_alignment,
             zone, p_throws, stand, inning_topbot, type, bb_type), 
           as.factor)
  ) %>%
  mutate(
    on_1b = ifelse(is.na(on_1b), 0, 1),
    on_2b = ifelse(is.na(on_2b), 0, 1),
    on_3b = ifelse(is.na(on_3b), 0, 1),
    events = fct_explicit_na(events, "no_event"),
    bb_type = fct_explicit_na(bb_type, "no_hit")
  )

# Create hitting and pitching subsets
batted_ball_df <- clean_df %>%
  filter(type == "X" & bb_type != "no_hit" & bb_type != "bunt")

top_pitches_list <- names(sort(table(clean_df$pitch_name), decreasing=TRUE)[1:8])
df_top_pitches <- clean_df %>%
  filter(pitch_name %in% top_pitches_list & pitch_name != "")
```

---

## 2. Visualizations

### 2.1 The Sweet Spot
Analyzing the relationship between Launch Angle and Launch Speed (Exit Velocity) to identify optimal contact profiles.

```{r sweet-spot}
s1_sweet_spot <- ggplot(batted_ball_df, aes(x = launch_angle, y = launch_speed)) +
  geom_point(alpha = 0.3, color = "blue", size = 1) + 
  geom_hline(yintercept = 95, color = "red", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = 8, color = "darkgreen", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = 32, color = "darkgreen", linetype = "dashed", linewidth = 1) +
  labs(
    title = "Figure: Launch Speed vs. Launch Angle",
    x = "Launch Angle (Degrees)",
    y = "Launch Speed (mph)",
    caption = "Red line = 95mph 'Hard Hit', Green lines = 8-32° 'Sweet Spot'"
  ) +
  theme_minimal()
s1_sweet_spot
```

### 2.2 Pitch Speed by Type
Comparing the pitch speed distributions of the top 8 most frequently used pitch types in the postseason.

```{r pitch-speed}
b1_pitch_speed <- ggplot(df_top_pitches, aes(x = pitch_name, y = release_speed)) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 1.5) +
  labs(
    title = "Figure: Boxplot of Release Speed by Pitch Type",
    x = "Pitch Type",
    y = "Release Speed (mph)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
b1_pitch_speed
```

---

## 3. Statistical Tests